Synthesizing Quotient Lenses

### 2018-03-13 papers/QuotientLenses/main.tex
merge

\begin{document}
\title{Synthesizing Quotient Lenses}
\begin{abstract}
{\em Quotient lenses} are bidirectional transformations that respect
specified equivalence relations with the goal of 
appropriately handling inessential details in data formats. 
For example, a programmer could use a quotient lens to define 
a transformation that ignores the order of fields in XML data, so
that two XML files with the same fields but in different orders would be
considered the same. 
This paper simplifies the task of programming quotient lenses in three
steps.
First, we introduce {\em quotient regular expressions} (QREs) to
make it easy to express which portions of a data format are
inessential.  A QRE is an augmented regular expression that
simultaneously defines a regular expression and an equivalence
relation over matching strings.  Variations within the same
equivalence class are deemed inessential. 
Next, we introduce {\em QRE lenses}, which are lenses that map
between QREs.
Our key technical result is a proof that every QRE lens can be
transformed into a QRE lens that has the form $c_1 \cdot \ell \cdot
c_2$, where $c_1$ and $c_2$ are {\em canonizers} 
and $\ell$ is a bijective transformation.  Intuitively,
the canonizers normalize the data  by picking a
canonical representative of the relevant equivalence relations. 
This theorem means we can push all the normalization to the edges of a
bidirectional transformation, leaving in the middle a pure bijective
transformation.  
Finally,
we levarage this insight to {\em synthesize} QRE lenses from a pair
of QREs and example input-output pairs by reusing earlier work on
synthesizing bijective lenses.
We have implemented QREs and QRE lens synthesis 
in the bidirectional programming language Boomerang.
We evaluate the effectiveness of our approach by using it to
synthesize QRE lenses between various real-world data formats in
the Optician benchmark suite.

@@

\begin{figure}[t]
\centering
\begin{subfigure}[b]{.49\textwidth}
\centering
\includegraphics{generated-graphs/times_opt}
\caption{}
\label{subfig:lenssize}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}
\includegraphics{generated-graphs/times_new.eps}
\caption{}
\label{subfig:examplesused}
\end{subfigure}
\caption{Runtimes of Optician and our system. In (a), we run Optician and our
  system on Optician's benchmarks. We find that there is only a negligible
  performance overhead incurred by using QREs. In (b), we run Optician with QREs
  on the Optician benchmarks we modified to use QREs. We find that it is able to
  synthesize all quotient lenses in under 20 seconds, and typically finished in
  under 5 seconds.
\label{fig:times}
\end{figure}

>>>>

\begin{document}
\title{Synthesizing Quotient Lenses}
\begin{abstract}
{\em Quotient lenses} are bidirectional transformations whose correctness
laws are ``loosened'' by specified equivalence relations, allowing
inessential details in concrete data formats to be suppressed. 
For example, a programmer could use a quotient lens to define 
a transformation that ignores the order of fields in XML data, so
that two XML files with the same fields but in different orders would be
considered the same, allowing a single, simple lens to handle them both.

This paper simplifies the task of programming quotient lenses with
three connected innovations.
First, we introduce {\em quotient regular expressions} (QREs), to
easily mark aspects of data formats that are
inessential.  A QRE is an augmented regular expression that
simultaneously defines a regular language and an equivalence
relation on this language.  Variations within the same
equivalence class are deemed inessential. 
Next, we introduce {\em QRE lenses}---lenses mapping
between QREs.
Our key technical result is a proof that every QRE lens can be
transformed into a QRE lens that has the form $c_1 \cdot \ell \cdot
c_2$, where $c_1$ and $c_2$ are {\em canonizers} 
and $\ell$ is a bijective transformation.  Intuitively,
the canonizers normalize data by picking a
canonical representative of the relevant equivalence relations. 
This theorem means we can push all normalization to the outer edges of a
bidirectional transformation, leaving in the middle a pure bijective
transformation.  
Finally,
we leverage this normalization theorem to {\em synthesize} QRE lenses from a pair
of QREs and example input-output pairs, reusing earlier work on
synthesizing plain bijective lenses.
We have implemented QREs and QRE lens synthesis 
in the bidirectional programming language Boomerang.
We evaluate the effectiveness of our approach by using it to
synthesize QRE lenses between various real-world data formats in
the Optician benchmark suite.

@@

\begin{figure}[t]
\centering
\begin{subfigure}[b]{.49\textwidth}
\centering
\includegraphics{generated-graphs/times_opt}
\caption{}
\label{subfig:lenssize}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}
\includegraphics{generated-graphs/times_new.eps}
\caption{}
\label{subfig:examplesused}
\end{subfigure}
\caption{Runtimes of Optician and our system. In (a), we run Optician and our
  system on Optician's benchmarks. We find that there is only a negligible
  performance overhead incurred by using QREs. In (b), we run Optician with QREs
  on the Optician benchmarks we modified to use QREs. We find that it is able to
  synthesize all quotient lenses in under 20 seconds, and typically finished in
  under 5 seconds.}
\label{fig:times}
\end{figure}
